---
title: "Non-stationary test"
author: "Zhe Sha"
date: "19 July 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loadfun}
library(sp)
library(INLA)
library(rgeos)
library(fields)
library(rgdal) 
```

## Simulate Non-stationary process

First generete the mesh and separate it into two regions.
```{r eg.partition}
set.seed(18)
## Define the regions
loc.bnd <- matrix(c(0,0, 5,0, 5,5, 0,5), 4, 2, byrow=TRUE)
segm.bnd <- inla.mesh.segment(loc.bnd)
out.Poly <- SpatialPolygons(list(Polygons(list(Polygon(loc.bnd)), ID = "out")))
loc.int <- matrix(c(0.87,0.27, 0.13,0.13, 0.5,0.35, 0.07,0.97, 
                    0.18,0.92, 0.31,0.32), byrow=T, 6, 2)*2 + 1.5
int.bound <- inla.nonconvex.hull(loc.int, convex=0.6)
int.Poly <- SpatialPolygons(list(Polygons(list(Polygon(int.bound$loc)), ID = "in")))
locPoly <- spsample(as(int.Poly, "SpatialLines"), n = 100, type = "regular")
segm.int <- inla.mesh.segment(locPoly)

## Sample dense and evenly spaced points in the entrie region
loc0 <- spsample(out.Poly, n = 500, type = "hexagonal", offset = c(0, 0))
## Remove points in the inner polygon
loc0in <- which(over(loc0, int.Poly) ==1)
loc0out <- loc0[-loc0in, ]
## Sample sparse and evenly spaced points in the inner polygon
loc1 <- spsample(int.Poly, n = 400, type = "hexagonal", offset = c(0, 0))

## combine in and out
locs <- rbind(loc0, loc1)
## Triangulation for the entire regions
mesh <- inla.mesh.2d(loc = locs, boundary = segm.bnd, 
                       interior = list(segm.int), offset = c(0.5,0.5),
                      cutoff = 0.05, max.edge = 0.5)
plot(mesh, asp =1)

## group the vertices
vsp <- SpatialPoints(mesh$loc[, 1:2])
Omega1 <- which(over(vsp, out.Poly) == 1)
Omega2 <- which(over(vsp, int.Poly) == 1)
boundsoutline <- as(out.Poly, "SpatialLines")
boundsout <- which(over(vsp, boundsoutline) == 1)
boundsin <- mesh$idx$segm

vOmega1 <- setdiff(setdiff(setdiff(Omega1, Omega2), boundsin),boundsout)
vOmega2 <- setdiff(Omega2, boundsin)
vboundsin <- boundsin[-c(1:4)]
vboundsout <- boundsout
vout <- which(is.na(over(vsp, out.Poly)))

## plot the points colored by group
plot(mesh, asp = 1)
points(vsp[vOmega1,], col = "blue")
points(vsp[vOmega2,], col = "red")
points(vsp[vboundsin,], pch = 19)
points(vsp[vboundsout,], pch = 19)
points(vsp[vout,], pch = 19, col = "blue")
```

Now based on the lable of the vertices, we build the weights for the non-stationary parameters 

```{r spdepars}
lsigma1 <- log(0.5)
lrho1 <- log(1)
lkappa1 <- log(8)/2 - lrho1
ltau1 <- 0.5*log(1/(4*pi)) - lsigma1 - lkappa1

lsigma2 <- log(5)
lrho2 <- log(0.2)
lkappa2 <- log(8)/2 - lrho2
ltau2 <- 0.5*log(1/(4*pi)) - lsigma2 - lkappa2

## define basis function for tau
nv <- nrow(mesh$loc)
np <- 5
Btmat <- matrix(0, nrow = np, ncol = nv)
## Omega1
bt1 <- c(ltau1, -1, 1, 0, 0)
## Omega2
bt2 <- c(ltau2, 0, 0, -1, 1)
## boundin
btb <- (bt1 + bt2)/2
Btmat[,c(vOmega1, vboundsout, vout)] <- bt1
Btmat[,vOmega2] <- bt2
Btmat[,vboundsin] <- btb
Btmat <- t(Btmat)

## define basis function for kappa
Bkmat <- matrix(0, nrow = np, ncol = nv)
## Omega1
bk1 <- c(lkappa1, 0, -1, 0, 0)
## Omega2
bk2 <- c(lkappa2, 0, 0, 0, -1)
## boundin
bkb <- (bk1 + bk2)/2
Bkmat[,c(vOmega1, vboundsout, vout)] <- bk1
Bkmat[,vOmega2] <- bk2
Bkmat[,vboundsin] <- bkb
Bkmat <- t(Bkmat)

## Stationary spde
spde0 <- inla.spde2.matern(mesh, 
                           B.tau = matrix(c(ltau1, -1, 1), 1, 3), 
                           B.kappa = matrix(c(lkappa1, 0, -1), 1,3),
                           theta.prior.mean = c(0,0), 
                           theta.prior.prec = c(1,1))
## Non-stationary spde
spde1 <- inla.spde2.matern(mesh, 
                           B.tau = Btmat, 
                           B.kappa =Bkmat,
                           theta.prior.mean = c(0,0,0,0), 
                           theta.prior.prec = c(1,1,1,1))

## Check the covariance structure
Q0 <- inla.spde.precision(spde0, c(0,0))
Q1 <- inla.spde.precision(spde1, c(0,0,0,0))
mvar0 <- sqrt(diag(inla.qinv(Q0)))
mvar1 <- sqrt(diag(inla.qinv(Q1)))

proj <- inla.mesh.projector(mesh, dims = c(100,100))

image.plot(proj$x, proj$y, inla.mesh.project(proj, as.vector(mvar0)), 
             col = topo.colors(24),
              xlab = "", ylab="", asp = 1)

image.plot(proj$x, proj$y, inla.mesh.project(proj, as.vector(mvar1)), 
             col = topo.colors(24),
              xlab = "", ylab="", asp = 1)


## Check simulated sample
xsamp <- rnorm(nrow(Q0))
Q0_chol <- chol(Q0)
Q1_chol <- chol(Q1)
tsamp0 <- as.vector(solve(t(Q0_chol), xsamp))
tsamp1 <- as.vector(solve(t(Q1_chol), xsamp))


image.plot(proj$x, proj$y, inla.mesh.project(proj, as.vector(tsamp0)), 
             col = topo.colors(24),
              xlab = "", ylab="", asp = 1)

image.plot(proj$x, proj$y, inla.mesh.project(proj, as.vector(tsamp1)), 
             col = topo.colors(24),
              xlab = "", ylab="", asp = 1)

```


## Inference on the non-stationary process 

The non-stationary construction looks all right. Next we try to infer the parameters and do posterior pridiction using simulated data.

First, we simulate the data for inference. We use the same SPDE as above but with different parameter values.
```{r simdata}
set.seed(10)
Qsim <- Q1 <- inla.spde.precision(spde1, c(0.5,0.2,-1,-0.1))
sampsim <- inla.qsample(1, Qsim)

dataloc <- expand.grid(proj$x, proj$y)
datapro <- inla.mesh.project(proj, as.vector(sampsim))
image.plot(proj$x, proj$y,datapro, 
             col = topo.colors(24),
              xlab = "", ylab="", asp = 1)

datasim <- data.frame(x=dataloc$Var1,  y = dataloc$Var2, z = as.vector(datapro))
datasim <- subset(datasim, x>=0 & x<=5 & y >=0 & y <=5)

```

Nest build the INLA model for the inference.
```{r inlainf}
data_loc <- as.matrix(datasim[,1:2])
obs <- datasim$z
  
Ay <- inla.spde.make.A(mesh = mesh, loc = data_loc)
st.est <- inla.stack(data = list(y=obs), A = list(Ay),
                     effects = list(process = 1:spde1$n.spde), tag = "est")
  
hyper <- list(prec = list(fixed = TRUE, initial = log(1)))
formula = y ~ -1 + f(process, model = spde1)
prec_scale <- 1/(0.001^2)
  
res_inla <- inla(formula, data = inla.stack.data(st.est, spde = spde1), family = "gaussian",
                  scale = prec_scale, control.family = list(hyper = hyper),
                   control.predictor=list(A=inla.stack.A(st.est), compute =TRUE))

res_pars <- inla.spde2.result(res_inla, "process", spde1, do.transf=TRUE)

## Plot the range 
plot(res_pars$marginals.range.nominal[[vOmega1[1]]], type = "l", main = "range 1")
plot(res_pars$marginals.range.nominal[[vOmega2[1]]], type = "l", main = "range 2")
plot(res_pars$marginals.range.nominal[[vboundsin[1]]], type = "l", main = "range boundary")
 
## Plot the variance
plot(res_pars$marginals.variance.nominal[[vOmega1[1]]], type = "l", main = "variance 1")
plot(res_pars$marginals.variance.nominal[[vOmega2[1]]], type = "l", main = "variance 2")
plot(res_pars$marginals.variance.nominal[[vboundsin[1]]], type = "l", main = "variance boundary")
 
```

The parameter looks all right. We also plot the posterior predictions.
```{r plotpost}
mean_post <- res_inla$summary.random$process$mean
sd_post <- res_inla$summary.random$process$sd

image.plot(proj$x, proj$y, inla.mesh.project(proj, as.vector(mean_post)), 
             col = topo.colors(24),
              xlab = "", ylab="", asp = 1)

image.plot(proj$x, proj$y, datapro, 
             col = topo.colors(24),
              xlab = "", ylab="", asp = 1)
```

