---
title: "Experiment 2a -- Update Mass -- use a different GRACE data set"
author: "Z Sha"
date: "1 Feb 2018"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)
## Math
#wd <-"/home/zs16444/coding/"
#dd <-"/home/zs16444/globalmass/"

# ## BS
wd <- "C:/ZSwork/"
dd <- "Z:/"

library(rgdal); library(sp);library(GEOmap)
library(INLA)
library(ncdf4)
library(gstat)

source(paste0(wd,"gmrcode/BHM_sphere/functions.R"))
source(paste0(wd, "gmrcode/BHM_sphere/partition_fun.R"))

## Genereate Fibonacci points on the sphere
fibo_points <- fiboSphere(N = 12960, L0 = TRUE)
fibo_points_xyz <- do.call(cbind, Lll2xyz(lat = fibo_points[,2], lon = fibo_points[,1]))
mesh0 <- inla.mesh.2d(loc = fibo_points_xyz, cutoff = 0.01, max.edge = 1)
## Make this "smoother"
mesh0 <- inla.mesh.2d(loc = mesh0$loc, cutoff = 0.01, max.edge = 1)
```

```{r loaddata, include=FALSE, message=FALSE, warning=FALSE, cache = TRUE}
## Load Maike's GRACE data
gdata <- read.table(paste0(dd, "WP2-SolidEarth/BHMinputs/GRACE/GRACE_v02.3b_trends_v03.txt"), header = T)
gm_loc <-  read.table(paste0(dd, "WP2-SolidEarth/BHMinputs/GRACE/GRACE_v02.3b_loc_v03.txt"), skip = 1)
n_grace <- ncol(gm_loc)
n_ll <- (n_grace-4)/2
names(gm_loc) <- c("id", "area", "lon_c", "lat_c", paste0(c("lon", "lat"), rep(1:n_ll, each = 2)))

## Create spatial polygons data frame
Polygon_list <- list()
for(i in 1:nrow(gm_loc)){
  lons <- na.omit(as.numeric(c(gm_loc[i, seq(5,ncol(gm_loc), 2)], gm_loc[i, 5])))
  lats <- na.omit(as.numeric(c(gm_loc[i, seq(6,ncol(gm_loc), 2)], gm_loc[i, 6])))
  Polygon_list[[i]] <- Polygon(cbind(lons, lats))
}

Polygons_list <- lapply(1:length(Polygon_list), function(x) Polygons(list(Polygon_list[[x]]), x))
SpPolygon <- SpatialPolygons(Polygons_list, proj4string = CRS("+proj=longlat"))

graceM <- SpatialPolygonsDataFrame(SpPolygon,gdata)
graceM$area <- gm_loc$area
gm_area <- geosphere::areaPolygon(graceM)/(1000^2) # km^2
graceM$area2 <- gm_area
graceM$diff <- graceM$area/graceM$area2
graceM$lon_c <- gm_loc$lon_c
graceM$lat_c <- gm_loc$lat_c

## Load the ice6g-ewh
ice6g <- read.table(paste0(dd, "WP2-SolidEarth/BHMinputs/GIA/GIA_Pel-6-VM5_ewh.txt"), header = T)
names(ice6g)[2:3] <- c("lon", "lat")
ice6g$lon <- ifelse(ice6g$lon <0, ice6g$lon + 360, ice6g$lon)
ice6g<- ice6g[order(ice6g$lat,ice6g$lon ),] # re-order the data according to the coordinates
coordinates(ice6g) <-c("lon", "lat")
gridded(ice6g) <- TRUE
proj4string(ice6g) <- CRS("+proj=longlat")
ice6g$areas <- geosphere::areaPolygon(as(ice6g, "SpatialPolygons"))/(1000^2)



## Load Rory's GRACE data
gr1 <- read.table(paste0(dd, "WP1-BHM/Experiment2a/grace_grid.txt"))
gr2 <-read.table(paste0(dd, "WP1-BHM/Experiment2a/grace-gia_grid.txt"))
lon <- 0.5:359.5
lat <- -89.5:89.5
locs <- expand.grid(lon, lat)
graceR <- data.frame(lon = locs[,1], lat = locs[,2], trend = gr1$V1, trendMgia = gr2$V1)
coordinates(graceR) <-c("lon", "lat")
gridded(graceR) <- TRUE
proj4string(graceR) <- CRS("+proj=longlat")
graceR$areas <- geosphere::areaPolygon(as(graceR, "SpatialPolygons"))/(1000^2)

# ## Output as netcdf
# graceout <- data.frame(lon = graceR@coords[,1], lat = graceR@coords[,2],
#                        trend= graceR@data$trend, trendMgia=graceR@data$trendMgia,
#                        std = graceR@data$std)
# 
# fname <- "GRACE_GSFC.nc"
# vars <- list(c("trend", "trend", "GRACE trend", "mm/yr"),
#              c("trendMgia", "trendMgia", "GRACE - GIA trend", "mm/yr"),
#              c("std", "std", "standard error sampled from MS", "mm/yr"))
# title <- "GRACE trend (GSFC)"
# 
# df2ncdf(graceout, fname, vars, title = title, append = FALSE)

## Add GRACE - i6g trend by hand
graceR$trendMgia2 <- graceR$trend - ice6g$trend
graceR$giadiff <- graceR$trendMgia - graceR$trendMgia2


## Load Ocean boundary
Ocean <- readOGR(dsn = paste0(dd, "WP1-BHM/maps/ne_110m_ocean"), layer = "ne_110m_ocean")
```

Since there was mass conservation problem in the GRACE super mascon data set, we now use a different GRACE data set to update mass.

Before using the BHM, we first load the previous data and new data for some intial checks.

# Check data
Check the integrated sum and mean of all the data. For the integrated mean, we define the following weights

1. global weight: area_i / Earth surface area

2. Ocean weight: area_i/ Ocean area

3. Land weight: area_i/Land area


## Super Mascons data

The area of the super mascons are provided in the original data set, however they are different from the area calculated by R using the definition of the polygons. The given areas of the polor polygons are roughly double of the calculated areas. See the plot below.

```{r maike_area, echo = FALSE}
spplot(graceM, "diff")
```

The total sum of the two versions of areas are 

```{r maike_area2, echo = FALSE}
cat("Give areas sum: ", sum(graceM$area),"\n",
    "Calculated areas sum: ", sum(graceM$area2))
```

The calculated area sum seems smaller. This might be caused by error in the definition of the polygons. In the following, we calculate the sum and mean using the areas given by MS.


First we calculate the sums.
```{r maike_sums, echo = FALSE}
## Find the polygons in Ocean
coords <- cbind(graceM$lon_c, graceM$lat_c)
coords[,1] <- ifelse(coords[,1] > 180, coords[,1]-360, coords[,1])
idxM <- unlist(over(Ocean, SpatialPoints(coords = coords, proj4string = CRS(proj4string(Ocean))), returnList = TRUE))

eartharea <- sum(graceM$area)
oceanarea <- sum(graceM$area[idxM])
landarea <-sum(graceM$area[-idxM])

Msum <- sum(graceM$mmweq*graceM$area)
Mocean <- sum(graceM$mmweq[idxM]*graceM$area[idxM])
Mland<- sum(graceM$mmweq[-idxM]*graceM$area[-idxM])

cat("The global sum of the super Mascons is:         ", Msum, "\n",
    "The sum of the super Mascons over the ocean is: ", Mocean, "\n",
    "The sum of the super Mascons over the land is:  ", Mland, "\n")
```

Next the we calculate the mean over the entire Earth surface.
```{r maike_means, echo = FALSE}
cat("Mean over the Earth \n",
"Global: ", Msum/eartharea, "\n",
"Ocean:  ", Mocean/eartharea, "\n",
"Land:   ", Mland/eartharea, "\n")
```

Next the we calculate the mean over the Ocean.
```{r maike_means2, echo = FALSE}
cat("Mean over the Ocean \n",
    "Global: ", Msum/oceanarea, "\n",
    "Ocean:  ", Mocean/oceanarea, "\n",
    "Land:   ", Mland/oceanarea, "\n")
```

Next the we calculate the mean over the Land.
```{r maike_means3, echo = FALSE}
landarea <- sum(graceM$area[-idxM])
cat("Mean over the Ocean \n",
"Global: ", Msum/landarea, "\n",
"Ocean:  ", Mocean/landarea, "\n",
"Land:   ", Mland/landarea, "\n")
```


## GIA data

We do the same sum and mean checks for the ice6g ewh data.

First we calculate the sums.
```{r gia_sums, echo = FALSE}
## Find the polygons in Ocean
coords <-ice6g@coords
coords[,1] <- ifelse(coords[,1] > 180, coords[,1]-360, coords[,1])
idxice <- unlist(over(Ocean, SpatialPoints(coords = coords, proj4string = CRS(proj4string(Ocean))), returnList = TRUE))

eartharea <- sum(ice6g$areas)
oceanarea <-sum(ice6g$areas[idxice])
landarea <- sum(ice6g$areas[-idxice])

giasum <- sum(ice6g$trend*ice6g$areas)
giaocean <- sum(ice6g$trend[idxice]*ice6g$areas[idxice])
gialand <-sum(ice6g$trend[-idxice]*ice6g$areas[-idxice])

cat("The global sum of the ice6g ewh is: ", giasum, "\n",
"The ocean sum of the ice6g ewh is:  ", giaocean, "\n",
"The land sum of the ice6g ewh is:   ", gialand, "\n")
```

Next the we calculate the mean over the entire Earth surface.
```{r gia_means, echo = FALSE}
cat("Mean over the Earth \n",
"Global: ", giasum/eartharea, "\n",
"Ocean:  ", giaocean/eartharea, "\n",
"Land:   ", gialand/eartharea, "\n")
```

Next the we calculate the mean over the Ocean.
```{r gia_means2, echo = FALSE}
cat("Mean over the Ocean \n",
"Global: ", giasum/oceanarea, "\n",
"Ocean:  ", giaocean/oceanarea, "\n",
"Land:   ", gialand/oceanarea, "\n")
```

Next the we calculate the mean over the Land.
```{r gia_means3, echo = FALSE}
cat("Mean over the Ocean \n",
"Global: ", giasum/landarea, "\n",
"Ocean:  ", giaocean/landarea, "\n",
"Land:   ", gialand/landarea, "\n")
```

### Compare MS GIA data and GSFC derived GIA

We compare the GIA in ewh to the one drived from GSFC by GRACE - (GRACE-GIA) data.
```{r giacompare, echo=FALSE}
ice6g$GSFCt <- graceR$trend - graceR$trendMgia
ice6g$diff <- ice6g$trend - ice6g$GSFCt
spplot(ice6g, c("trend", "GSFCt"))
spplot(ice6g, "diff")
```


## New GRACE data

Also check for the new GRACE data (GSFC provided by RB).

First we calculate the sums.
```{r rory_sums, echo = FALSE}
## Find the polygons in Ocean
coords <-graceR@coords
coords[,1] <- ifelse(coords[,1] > 180, coords[,1]-360, coords[,1])
idxR <- unlist(over(Ocean, SpatialPoints(coords = coords, proj4string = CRS(proj4string(Ocean))), returnList = TRUE))

eartharea <- sum(graceR$areas)
oceanarea <-sum(graceR$areas[idxR])
landarea <- sum(graceR$areas[-idxR])

trendsum1 <- sum(graceR$trend*graceR$areas)
trendocean1 <- sum(graceR$trend[idxR]*graceR$areas[idxR])
trendland1 <-sum(graceR$trend[-idxR]*graceR$areas[-idxR])

cat("Sum for the GRACE trend \n",
"The global sum of the GRACE trend is: ", trendsum1, "\n",
"The ocean sum of the GRACE trend is:  ", trendocean1, "\n",
"The land sum of the GRACE trend is:   ", trendland1, "\n")
```

Next the we calculate the mean over the entire Earth surface.
```{r rory_means, echo = FALSE}
cat("GRACE trend mean over the Earth \n",
"Global: ", trendsum1/eartharea, " (Rory's: -0.000907213)", "\n",
"Ocean:  ", trendocean1/eartharea, "\n",
"Land:   ", trendland1/eartharea, "\n")
```

Next the we calculate the mean over the Ocean.
```{r rory_means2, echo = FALSE}
cat("GRACE trend mean over the ocean \n",
"Global: ", trendsum1/oceanarea, " (Rory's:  -0.001276672)","\n",
"Ocean:  ", trendocean1/oceanarea, "         ( 0.6895327)", "\n",
"Land:   ", trendland1/oceanarea, "        (-0.6908094)", "\n")
```

Next the we calculate the mean over the Land.
```{r rory_means3, echo = FALSE}
cat("GRACE trend mean over the land \n",
"Global: ", trendsum1/landarea, "\n",
"Ocean:  ", trendocean1/landarea, "\n",
"Land:   ", trendland1/landarea, "( -1.696294)", "\n")
```

## GRACE - GIA (ice6g)

The GRACE - GIA(ice6g) trend is also provided by RB from the GSFC data. We will use this data to update the mass trend later. First We compare this result with GRACE(GSFC) - ice6g(MS). 

We calculate the sums for both.
```{r gg_sums, echo = FALSE}
## Find the polygons in Ocean
coords <-graceR@coords
coords[,1] <- ifelse(coords[,1] > 180, coords[,1]-360, coords[,1])
idxR <- unlist(over(Ocean, SpatialPoints(coords = coords, proj4string = CRS(proj4string(Ocean))), returnList = TRUE))

eartharea <- sum(graceR$areas)
oceanarea <-sum(graceR$areas[idxR])
landarea <- sum(graceR$areas[-idxR])

trendsum1 <- sum(graceR$trendMgia*graceR$areas)
trendocean1 <- sum(graceR$trendMgia[idxR]*graceR$areas[idxR])
trendland1 <-sum(graceR$trendMgia[-idxR]*graceR$areas[-idxR])

trendsum2 <- sum(graceR$trendMgia2*graceR$areas)
trendocean2 <- sum(graceR$trendMgia2[idxR]*graceR$areas[idxR])
trendland2 <-sum(graceR$trendMgia2[-idxR]*graceR$areas[-idxR])

cat("Sum for the trend RB  \n",
"The global sum is: ", trendsum1, "\n",
"The ocean sum is:   ", trendocean1, "\n",
"The land sum is:   ", trendland1, "\n")

cat("Sum for the trend RB-MS \n",
"The global sum is:  ", trendsum2, "\n",
"The ocean sum is:   ", trendocean2, "\n",
"The land sum is:   ", trendland2, "\n")
```

Next the we calculate the mean over the entire Earth surface.
```{r gg_means, echo = FALSE}
cat("RB trend over the Earth \n",
"Global: ", trendsum1/eartharea, " (Rory's: 0.001688293)", "\n",
"Ocean:  ", trendocean1/eartharea, "\n",
"Land:   ", trendland1/eartharea, "\n")

cat("\n RB - MS trend over the Earth \n",
"Global: ", trendsum2/eartharea, "\n",
"Ocean:  ", trendocean2/eartharea, "\n",
"Land:   ", trendland2/eartharea, "\n")
```

Next the we calculate the mean over the Ocean.
```{r gg_means2, echo = FALSE}
cat("RB trend over the ocean \n",
"Global: ", trendsum1/oceanarea, " (Rory's:  0.002375843)","\n",
"Ocean:  ", trendocean1/oceanarea, "         ( 1.118967)", "\n",
"Land:   ", trendland1/oceanarea, "        (-1.116591)", "\n")
 
cat("\n RB - MS trend  over the ocean \n",
"Global: ", trendsum2/oceanarea,  "\n",
"Ocean:  ", trendocean2/oceanarea, "\n",
"Land:   ", trendland2/oceanarea, "\n")
```

Next the we calculate the mean over the Land.
```{r gg_means3, echo = FALSE}
cat("RB trend mean over the land \n",
"Global: ", trendsum1/landarea, "\n",
"Ocean:  ", trendocean1/landarea, "\n",
"Land:   ", trendland1/landarea, "( -2.741808)", "\n")

cat("\n RB - MS trend over the land \n",
"Global: ", trendsum2/landarea, "\n",
"Ocean:  ", trendocean2/landarea, "\n",
"Land:   ", trendland2/landarea,  "\n")
```

## Find error estimates for the new GRACE data

The is no error estimates in the new GRACE data, so we try to derive it from the GRACE data provided by MS. Below shows the standard error of MS' data.

```{r gerrorplot, echo=FALSE}
spplot(graceM, "std")
```

We sample the error from this data to the 1 degree grid.
```{r gerror, echo=FALSE}
graceR$std <- over(graceR, graceM)$std
spplot(graceR, "std")
```


# BHM Analysis

## Prior setup

Use the polygon data to have rough estimate of the correlation length and variance.
```{r Grace_prior, echo = FALSE}
#v2 <- variogram(trendMgia~1, graceR[sample(1:nrow(graceR), 3000),]) 
#plot(v2)

## Priors mean and variance for the parameters: rho and sigma
mu_r <- 2500/6371
v_r <- 1
mu_s <- 25 # area scale to about 100km^2 ~ 1 degree resolution
v_s <- 50^2

## Transform the parameters for the SPDE_GMRF approximation
trho <- Tlognorm(mu_r, v_r)
tsigma <- Tlognorm(mu_s, v_s)

## Build the SPDE model with the prior
lsigma0 <- tsigma[1]
theta1_s <- tsigma[2]
lrho0 <- trho[1]
theta2_s <- trho[2]
lkappa0 <- log(8)/2 - lrho0
ltau0 <- 0.5*log(1/(4*pi)) - lsigma0 - lkappa0

M_spde <- inla.spde2.matern(mesh0, B.tau = matrix(c(ltau0, -1, 1),1,3), B.kappa = matrix(c(lkappa0, 0, -1), 1,3),
                              theta.prior.mean = c(0,0), theta.prior.prec = c(sqrt(1/theta1_s), sqrt(1/theta2_s)))

```


## Generate mesh 

We also need a mesh to represent the mass process. For the same reason as discussed in previous section for SSH, we need a mesh with 1 degree resolution. The GRACE data is all over the globe, so we can use the \code{mesh0}.

## Link data to process
```{r grace_area, cache = TRUE}
grace_loc <- do.call(cbind, Lll2xyz(lon = graceR@coords[,1], lat = graceR@coords[,2]))
A_GRACE_data <- inla.spde.make.A(mesh = mesh0, loc = grace_loc)
A_M_pred <- inla.spde.make.A(mesh = mesh0, loc = rbind(mesh0$loc))
```

We use the GIA corrected GRACE trend in the data.
```{r stacks, eval = FALSE}
## Create the estimation and prediction stack
st.est <- inla.stack(data = list(y=graceR$trendMgia), A = list(A_GRACE_data),
                     effects = list(M= 1:M_spde$n.spde), tag = "est")
st.pred <- inla.stack(data = list(y=NA), A = list(rbind(A_GRACE_data, A_M_pred)),
                      effects = list(M=1:M_spde$n.spde), tag = "pred")
stM <- inla.stack(st.est, st.pred)
```

## INLA inference

Now we can run INLA for the Bayesian inference. Do not run on a desktop the process may use up to 64GB memory at peak. We ran this on a server with enough memory.

```{r inla_run_grace, include = TRUE, eval = FALSE}
## Fix altimetry errors as they are known
hyper <- list(prec = list(fixed = TRUE, initial = 0))
prec_scale <- c(1/graceR$std^2, rep(1, nrow(A_GRACE_data) + nrow(A_M_pred)))

## The formular for modelling the SSH mean
formula = y ~ -1 +  f(M, model = M_spde)

## Run INLA
res_inla <- inla(formula, data = inla.stack.data(stM, spde = M_spde), family = "gaussian",
                 scale =prec_scale, control.family = list(hyper = hyper),
                 control.predictor=list(A=inla.stack.A(stM), compute =TRUE))
```


## Results
### Assemble and save results

Aftern running the INLA procedure, we assemble the inference and prediction results and save them as "exp2a_M2.RData".
```{r inla_res_grace, include = TRUE, eval = FALSE, echo = FALSE}
INLA_pred <- res_inla$summary.linear.predictor
## Extract and project predictions
pred_idx <- inla.stack.index(stM, tag = "pred")$data[1:64800]


## mass
M_m <- INLA_pred$mean[pred_idx] 
M_u <- INLA_pred$sd[pred_idx]
proj <- inla.mesh.projector(mesh0, projection = "longlat", dims = c(360,180), xlim = c(0,359), ylim = c(-89.5, 89.5))
M_grid <- expand.grid(proj$x, proj$y)
M_pred <- data.frame(lon = M_grid[,1], lat = M_grid[,2],
                       mean = M_m, u = M_u, diff = M_m - graceR@data$trendMgia)

res_M2 <- list(res_inla = res_inla, spde = M_spde, st = stM, 
            mesh = mesh0,  M_pred = M_pred, Adata = A_GRACE_data, Apred = A_M_pred)

graceR@data$pred_mean <- M_m
graceR@data$pred_u <- M_u
graceR@data$pred_diff <- M_m - graceR@data$trendMgia

save(res_M2, grace, file =paste0(dd, "WP1-BHM/Experiment2a/exp2a_M2.RData"))
```


### Plot the posteriors of the hyper parameters

```{r hyper_grace, include=TRUE, echo=FALSE}
load(paste0(dd, "WP1-BHM/Experiment2a/exp2a_M2zero.RData"))
pars_M <- marginal_par(res = res_M2, process = "M", plot = TRUE)
## The posterior modes
print(paste("The estimated correlation lengths are:", pars_M$rho_mode*6371,  sep = "  "))

print(paste("The estimated marginal variances are:", pars_M$sigma_mode,sep = "  "))
```


## Plot the predictions

We first plot the predicted $X_{mass}$ on a 1 degree grid and then the aggregated $X_m$ on the same super mascons grid as the GRACE data.
```{r predict_grace, include=TRUE, echo=FALSE}
M_pred <- res_M2$M_pred

M_pred$mean2 <- ifelse(abs(M_pred$mean) > 19, sign(M_pred$mean)*20, M_pred$mean)
M_pred$u2 <- ifelse(abs(M_pred$u) > 5, 7, M_pred$u)
## plot the mean 
lattice::levelplot(mean2 ~ lon + lat, data = M_pred, aspect = "iso", at = seq(-20, 20, 2),
                     panel = function(x,y,z,...){
                       lattice::panel.levelplot(x,y,z,...)
                       map2 <- map("world2", interior = FALSE, plot = FALSE)
                       lattice::panel.xyplot(x=map2$x, y=map2$y, type = "l", col = "black")
                     },
                     main = "The predicited mass change  (mm/yr ewh)", xlab = "longitude", ylab = "latitude")

## plot the uncertainty
lattice::levelplot(u2 ~ lon + lat, data = M_pred, aspect = "iso", col.regions = topo.colors(20),
                     panel = function(x,y,z,...){
                       lattice::panel.levelplot(x,y,z,...)
                       map2 <- map("world2", interior = FALSE, plot = FALSE)
                       lattice::panel.xyplot(x=map2$x, y=map2$y, type = "l", col = "black")
                     },
                     main = "The predicted uncertainties  (mm/yr ewh)", xlab = "longitude", ylab = "latitude")
```

# Sanity checks on the updated mass

To check whether the updated mass is reasonable, we calculate the mass trend averaged over Earth sphere, ocean, and land.
```{r globalmean, echo=FALSE}
coords <-graceR@coords
coords[,1] <- ifelse(coords[,1] > 180, coords[,1]-360, coords[,1])
idxR <- unlist(over(Ocean, SpatialPoints(coords = coords, proj4string = CRS(proj4string(Ocean))), returnList = TRUE))

eartharea <- sum(graceR$areas)
oceanarea <-sum(graceR$areas[idxR])
landarea <- sum(graceR$areas[-idxR])

predsum <- sum(graceR$pred_mean*graceR$areas)
predocean <- sum(graceR$pred_mean[idxR]*graceR$areas[idxR])
predland <-sum(graceR$pred_mean[-idxR]*graceR$areas[-idxR])

```


```{r oceanmean, echo = FALSE}
cat("GRACE predicted trend mean over the Earth \n",
"Global: ", predsum/eartharea, "\n",
"Ocean:  ", predocean/eartharea, "\n",
"Land:   ", predland/eartharea, "\n")

cat("GRACE predicted trend mean over the ocean \n",
"Global: ", predsum/oceanarea, "\n",
"Ocean:  ", predocean/oceanarea, "\n",
"Land:   ", predland/oceanarea, "\n")

cat("GRACE predicted trend mean over the land \n",
"Global: ", predsum/landarea, "\n",
"Ocean:  ", predocean/landarea, "\n",
"Land:   ", predland/landarea, "\n")

```

