---
title: "Experiment 2C uing large foot print GRACE data and SPDE non-stationary setting"
author: "Z Sha"
date: "5 June 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
##module load /opt/modulefiles/gdal-2.2.3-x86_64
if(Sys.info()["nodename"] == "IT034844"){
  ## BS
  wd <- "C:/ZSwork/"
  dd <- "Z:/"
}else if(Sys.info()["nodename"] == "it064613"){
  ## Maths
  wd <-"/home/zs16444/coding/"
  dd <-"/home/zs16444/globalmass/"
}else{
  ## any server  
  wd <- "~/"
  dd <- "/./projects/GlobalMass/"
}

library(rgdal); library(sp);library(GEOmap)
library(INLA)
library(ncdf4)
source(paste0(wd,"gmrcode/BHM_sphere/functions.R"))
source(paste0(wd, "gmrcode/BHM_sphere/partition_fun.R"))
gracedata = "masscons"
```

## Introduction

We do the same experiment as previous but in this one, we use a new GRACE data that use large catchment reigon on land and large mascon polygons in the ocean. These spatial units are more natural for the GRACE and mass signals.

## The new GRACE data

Here we show the new GRACE data. The raw GRACE data are equal-areal mascons. We aggregated the raw data to the desired spatial unit. We have also taken care of the correlation in the measuremnt errors by inflating the errors.

```{r grace}
load(paste0(dd, "WP1-BHM/Experiment2c/data/grace_sp.rda"))

## Scale the data for plot
Land_spf$trend2 <- ifelse(abs(Land_spf$trend) > 19, sign(Land_spf$trend)*20, Land_spf$trend)
colpal1 <- rev(c('#b2182b','#d6604d','#f4a582','#fddbc7','#f7f7f7','#d1e5f0','#92c5de','#4393c3','#2166ac'))
spplot(Land_spf, "trend2", at =c(-20, -15, -10, -5, 0, 5, 10, 15, 21), col.regions = colpal1)

Land_spf$std3 <- ifelse(abs(Land_spf$std2) > 2.5, 2.9, Land_spf$std2)
colpal2 <- c('#fff5eb','#fee6ce','#fdd0a2','#fdae6b','#fd8d3c','#f16913','#d94801','#8c2d04')
spplot(Land_spf, "std3", at = c(0, 0.5, 1, 1.5, 2, 2.5, 3), col.regions = colpal2 )


Ocean_spf$trend2 <- ifelse(abs(Ocean_spf$trend) > 19, sign(Ocean_spf$trend)*20, Ocean_spf$trend)
spplot(Ocean_spf, "trend2", at =c(-20, -15, -10, -5, 0, 5, 10, 15, 21), col.regions = colpal1)

Ocean_spf$std3 <- ifelse(abs(Ocean_spf$std2) > 2.5, 2.9, Ocean_spf$std2)
spplot(Ocean_spf, "std3", at = c(0, 0.5, 1, 1.5, 2, 2.5, 3), col.regions = colpal2)

Greenland$trend2 <- ifelse(abs(Greenland$trend) > 19, sign(Greenland$trend)*20,Greenland$trend)
spplot(Greenland, "trend2", at =c(-20, -15, -10, -5, 0, 5, 10, 15, 21), col.regions = colpal1)

Greenland$std3 <- ifelse(Greenland$std2 > 2.5, 2.9, Greenland$std2)
spplot(Greenland, "std3", at = c(0, 0.5, 1, 1.5, 2, 2.5, 3), col.regions = colpal2 )

Antarctica$trend2 <-  ifelse(abs(Antarctica$trend) > 19, sign(Antarctica$trend)*20, Antarctica$trend)
spplot(Antarctica, "trend2", at =c(-20, -15, -10, -5, 0, 5, 10, 15, 21), col.regions = colpal1)

Antarctica$std3 <- ifelse(Antarctica$std2 > 2.5, 2.9, Antarctica$std2)
spplot(Antarctica, "std3", at = c(0, 0.5, 1, 1.5, 2, 2.5, 3), col.regions = colpal2 )

```

## Link the process to data

```{r mesh, include=FALSE}
## Load the same setting of data and mesh as in Experiment 2c_a
source(paste0(wd, "experiments/Experiment2/Doc/c/Rscripts/exp2c_loadData_nn.R"))
source(paste0(wd, "experiments/Experiment2/Doc/c/Rscripts/exp2c_INLAstacks_nn.R"))

```


# INLA inference

Now we can run INLA for the Bayesian inference. Do not run on a desktop the process may use up to 64GB memory at peak. We ran this on a server with enough memory.

```{r inla_run, include = TRUE, eval = FALSE}
## Fix altimetry errors as they are known
hyper <- list(prec = list(fixed = TRUE, initial = 0))

## The formular -- we add the constraint that mass change sum to zero
## constraint 1 -- vertices sum to zero
A1 <- matrix(1, nrow = 1, ncol = mesh0$n)
## Constraint 2 -- grace loc sum to zero
A2 <- matrix(colSums(A_grace), nrow = 1)
## Constraint 3 -- longlat grid sum to zero
gridll <- ice6g_vlm@coords
gridxyz <- do.call(cbind, Lll2xyz(lat = gridll[,2], lon = gridll[,1]))
A_grid <- inla.spde.make.A(mesh = mesh0, loc = gridxyz)
weights <- ice6g_vlm$areas/sum(ice6g_vlm$areas)
A3 <- matrix(weights, nrow = 1) %*% A_grid
A <- as.matrix(rbind(A1, A2, A3))
## The formular -- we add the constraint that mass change sum to zero
formula = y ~ -1 + f(mass, model = mass_spde, extraconstr = list(A = A, e = c(0, 0, 0))) +  f(steric, model = steric_spde)

## Run INLA
## No GIA error
res_inla <- inla(formula, data = inla.stack.data(stkall), family = "gaussian",
                 scale =prec_scale, 
                 control.family = list(hyper = hyper), control.compute = list(config = TRUE),
                 control.predictor=list(A=inla.stack.A(stkall), compute = TRUE), 
                 verbose = TRUE)
saveRDS(res_inla, file = "/./projects/GlobalMass/WP1-BHM/Experiment2c/Exp2c_a_nn.rds")

## With GIA error
res_inla2 <- inla(formula, data = inla.stack.data(stkall), family = "gaussian",
                 scale =prec_scale2, 
                 control.family = list(hyper = hyper), control.compute = list(config = TRUE),
                 control.predictor=list(A=inla.stack.A(stkall), compute = TRUE), 
                 verbose = TRUE)
saveRDS(res_inla2, file = "/./projects/GlobalMass/WP1-BHM/Experiment2c/Exp2c_a_lg2.rds")

```

# Results

## Assemble and save results

Now assemble the inla inference and prediction results for the result from \code{formular2}.
```{r inla_res, include = TRUE, eval = FALSE, echo = FALSE}
source(paste0(wd, "experiments/Experiment2/Doc/c/Rscripts/exp2c_a_resnn.R"))
```


## Plot the posteriors of the hyper parameters

```{r hyper, include=TRUE, echo = FALSE}
load(paste0(dd, "WP1-BHM/Experiment2c/Exp2c_a_lgnn.RData"))
## The posterior modes
print(paste("The estimated correlation length for mass in Ocean is:", pars_massO$rho_mode*6371, "km", sep = " "))
print(paste("The estimated marginal variance for mass in Ocean is:", sqrt(pars_massO$sigma_mode), "mm/yr", sep = " "))

print(paste("The estimated correlation length for mass in Land is:", pars_massL$rho_mode*6371, "km", sep = " "))
print(paste("The estimated marginal variance for mass in Land is:", sqrt(pars_massL$sigma_mode), "mm/yr", sep = " "))

print(paste("The estimated correlation length for steric is:", pars_steric$rho_mode*6371, "km", sep = "  "))
print(paste("The estimated marginal variance for steric is:", sqrt(pars_steric$sigma_mode), "mm/yr", sep = "  "))

```


## Plot the predictions

```{r predict, include=TRUE, echo = FALSE}
steric_pred <- ress_2c_a$pred$steric
ssh_pred <- ress_2c_a$pred$ssh
mass_pred <- ress_2c_a$pred$mass

## plot the mass mean 
lattice::levelplot(mean2 ~ lon + lat, data = mass_pred, aspect = "iso", at = seq(-20, 20, 2),
                     panel = function(x,y,z,...){
                       lattice::panel.fill(col = "grey")
                       lattice::panel.levelplot(x,y,z,...)
                       map2 <- map("world2", interior = FALSE, plot = FALSE)
                       lattice::panel.xyplot(x=map2$x, y=map2$y, type = "l", col = "black")
                     },
                     main = "The predicted mass trend (mm/yr ewh)", xlab = "longitude", ylab = "latitude")

## plot the mass uncertainty
lattice::levelplot(u2 ~ lon + lat, data = mass_pred, aspect = "iso", at = seq(0, 4, 0.5),col.regions = topo.colors(10),
                     panel = function(x,y,z,...){
                       lattice::panel.fill(col = "grey")
                       lattice::panel.levelplot(x,y,z,...)
                       map2 <- map("world2", interior = FALSE, plot = FALSE)
                       lattice::panel.xyplot(x=map2$x, y=map2$y, type = "l", col = "black")
                     },
                     main = "The predited mass uncertainties (mm/yr ewh)", xlab = "longitude", ylab = "latitude")

## Plot the steric mean
lattice::levelplot(mean2 ~ lon + lat, data = steric_pred, aspect = "iso", at = seq(-20, 20, 2),
                     panel = function(x,y,z,...){
                       lattice::panel.fill(col = "grey")
                       lattice::panel.levelplot(x,y,z,...)
                       map2 <- map("world2", interior = FALSE, plot = FALSE)
                       lattice::panel.xyplot(x=map2$x, y=map2$y, type = "l", col = "black")
                     },
                     main = "The predicted steric trend (mm/yr)", xlab = "longitude", ylab = "latitude")

## plot the steric uncertainty
lattice::levelplot(u2 ~ lon + lat, data = steric_pred, aspect = "iso", at = seq(0, 4, 0.5),col.regions = topo.colors(10),
                     panel = function(x,y,z,...){
                       lattice::panel.fill(col = "grey")
                       lattice::panel.levelplot(x,y,z,...)
                       map2 <- map("world2", interior = FALSE, plot = FALSE)
                       lattice::panel.xyplot(x=map2$x, y=map2$y, type = "l", col = "black")
                     },
                     main = "The predited steric uncertainties (mm/yr)", xlab = "longitude", ylab = "latitude")

## plot the ssh mean 
lattice::levelplot(mean2 ~ lon + lat, data = ssh_pred, aspect = "iso", at = seq(-20, 20, 1),
                     panel = function(x,y,z,...){
                       lattice::panel.fill(col = "grey")
                       lattice::panel.levelplot(x,y,z,...)
                       map2 <- map("world2", interior = FALSE, plot = FALSE)
                       lattice::panel.xyplot(x=map2$x, y=map2$y, type = "l", col = "black")
                     },
                     main = "The predicted SSH trend (mm/yr)", xlab = "longitude", ylab = "latitude")

## plot the ssh uncertainty
lattice::levelplot(u2 ~ lon + lat, data = ssh_pred, aspect = "iso", at = seq(0, 4, 0.5),col.regions = topo.colors(10),
                     panel = function(x,y,z,...){
                       lattice::panel.fill(col = "grey")
                       lattice::panel.levelplot(x,y,z,...)
                       map2 <- map("world2", interior = FALSE, plot = FALSE)
                       lattice::panel.xyplot(x=map2$x, y=map2$y, type = "l", col = "black")
                     },
                     main = "The predited SSH uncertainties (mm/yr)", xlab = "longitude", ylab = "latitude")
```


```{r gracepred}
graceO_sp <- ress_2c_a$pred$graceO
graceO_sp$diff <- graceO_sp$predmean - graceO_sp$trend
spplot(graceO_sp, "predmean")


graceL_sp <- ress_2c_a$pred$graceL
#graceL_sp$aa <- as.numeric(aa)
graceL_sp$predmean2 <- ifelse(abs(graceL_sp$predmean) > 19, sign(graceL_sp$predmean)*20, graceL_sp$predmean)
graceL_sp$diff <- graceL_sp$predmean - graceL_sp$trend
spplot(graceL_sp, "diff", at =c(-20, -15, -10, -5, 0, 5, 10, 15, 20), col.regions = colpal1)
spplot(graceL_sp, c("predmean2"),  at =c(-20, -15, -10, -5, 0, 5, 10, 15, 20), col.regions = colpal1)

grace_green <- ress_2c_a$pred$graceGL
grace_green$diff <- grace_green$predmean - grace_green$trend
spplot(grace_green, c("predmean", "trend"))
spplot(grace_green, "diff")

grace_ant <- ress_2c_a$pred$graceAnt
grace_ant$diff <- grace_ant$predmean - grace_ant$trend
spplot(grace_ant, c("predmean", "trend"))
spplot(grace_ant, "diff")
```



# Predict steric on Basin level 
To compare our prediction with other existing work, we also predict the averge steric change at the the basin level. This can be done either before or after calling the INLA procedure. Here we use both and also compare the results. The uncertainties will be different due to non-zero correlation.

```{r basin1}
load(paste0(dd, "WP1-BHM/maps/Ocean/basins.rda"))
pred_basin <- ress_2c_a$pred$basins

allbasinsdf$trend1 <- pred_basin$mean
allbasinsdf$trend2 <- pred_basin$meanpost
allbasinsdf$u1 <- pred_basin$u
allbasinsdf$u2 <- pred_basin$u2
allbasinsdf$u3 <- pred_basin$upost

spplot(allbasinsdf, "trend1", at = c(-2.5, -1.5, -0.5, 0.5, 1.5, 2.5, 3.5, 4.5), col.regions = rev(c('#d73027','#f46d43','#fdae61','#fee090','#ffffbf','#e0f3f8','#abd9e9','#74add1','#4575b4')),
       main = "Basin trend (posterior mean)")

spplot(allbasinsdf, "trend2", at = c(-2.5, -1.5, -0.5, 0.5, 1.5, 2.5, 3.5, 4.5), col.regions = rev(c('#d73027','#f46d43','#fdae61','#fee090','#ffffbf','#e0f3f8','#abd9e9','#74add1','#4575b4')),
       main = "Basin trend (posterior sample mean)")

spplot(allbasinsdf, "u1", at = c(0, 0.04, 0.08, 0.12, 0.16, 0.2), 
       col.regions = c('#fef0d9','#fdd49e','#fdbb84','#fc8d59','#ef6548','#d7301f','#990000'),
       main = "uncertainty (posterior covariance approximation)")

spplot(allbasinsdf, "u2", at = c(0, 0.04, 0.08, 0.12, 0.16, 0.2), 
       col.regions = c('#fef0d9','#fdd49e','#fdbb84','#fc8d59','#ef6548','#d7301f','#990000'),
       main = "uncertainty (independent errors)")

spplot(allbasinsdf, "u3", at = c(0, 0.04, 0.08, 0.12, 0.16, 0.2), 
       col.regions = c('#fef0d9','#fdd49e','#fdbb84','#fc8d59','#ef6548','#d7301f','#990000'),
       main = "uncertainty (posterior sample approximation)")


## compute the global steric
pred_basin2 <- data.frame(names = c("SAt", "ISPac", "EPac", "STNAt", "SPNAt", "NWPac", "global"), 
                          mean = pred_basin$mean, sd=pred_basin$u, model = "BHM_catchment")
write.csv(pred_basin2, file="~/bhm_basin2.csv")

```